---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---
## Loading and preprocessing the data
```{r chunk1}
library(ggplot2)
library(data.table)
library(DMwR)

dat <- data.table(read.csv("activity.csv"))
```

## What is mean total number of steps taken per day?
```{r chunk 2}
NumOfDays <- length(unique(dat$date))
SumOfEachDay <- tapply(dat$steps, dat$date, sum, na.rm = TRUE)
MeanTotalSteps <- mean(SumOfEachDay)
MedianTotalSteps <- median(SumOfEachDay)
hist(x = SumOfEachDay)
MeanTotalSteps
MedianTotalSteps
```

## What is the average daily activity pattern?
```{r chunk3}
IntMean <- tapply(dat$steps, dat$interval, mean, na.rm = TRUE)
plot(IntMean, type = "l", xlab = "5-Minute-Interval", ylab = "Mean Steps taken per day", main = "Mean steps taken per day in each interval")
MaxStepsInt <- which.max(IntMean)
MaxStepsInt
```

## Imputing missing values
In this section we use the K-Neares-Neighbour approach to impute the missing values. For this we make use of the DMwR package, which is available on CRAN.
```{r chunk4}
LogicalNA <- is.na(dat)
NumOfNA <- sum(LogicalNA)
datImp <- knnImputation(dat, k = 10)

ImpSumOfEachDay <- tapply(datImp$steps, datImp$date, sum)
ImpMeanTotalSteps <- mean(ImpSumOfEachDay)
ImpMedianTotalSteps <- median(ImpSumOfEachDay)
hist(x = ImpSumOfEachDay)
ImpMeanTotalSteps
ImpMedianTotalSteps
```


## Are there differences in activity patterns between weekdays and weekends?
```{r chunk5}
datImp2 <- cbind(datImp, weekdays(as.Date(datImp$date)))
names(datImp2) <- c("steps", "date", "interval", "weekday")
datImp2 <- data.table(datImp2)
datImp2[grepl(pattern = "Montag|Dienstag|Mittwoch|Donnerstag|Freitag", x = `weekday`), "weekday or weekend"] <- "weekday"
datImp2[grepl(pattern = "Samstag|Sonntag", x = `weekday`), "weekday or weekend"] <- "weekend"
datImp2[, `weekday or weekend` := as.factor(`weekday or weekend`)]

ggplot(datImp2 , aes(x = interval , y = steps, color=`weekday or weekend`)) + geom_line() + labs(title = "Mean Steps by Weekday/-end", x = "Interval", y = "Number of Steps") + facet_wrap(~`weekday or weekend` , ncol = 1, nrow=2)
``` 